---
title: "Global Vaccination & Science Attitudes"
format: 
  dashboard:
    theme: cosmo
    orientation: columns
    nav-tabs: true
    nav-buttons: 
      - icon: github
        href: https://github.com/LiamQuach/
        aria-label: GitHub
      - icon: linkedin
        href: https://www.linkedin.com/in/liam-quach-36a52a197/
        aria-label: LinkedIn
      - icon: twitter
        href: https://twitter.com/username/
        aria-label: Twitter
      - icon: instagram
        href: https://instagram.com/username/
        aria-label: Instagram
      - icon: envelope
        href: mailto:email@example.com
        aria-label: Email
logo: images/welcome_logo.jpg
logo-alt: Welcome logo with big W
---

```{r}
#| label: setup
#| include: false

# Setup: Load Libraries and Data
library(tidyverse)
library(readxl)
library(leaflet)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(patchwork)
library(plotly)
library(htmltools)
library(flexdashboard)

# Create SVG logo (Olympic rings style) - Save this in images directory
create_vax_logo <- function() {
  if (!dir.exists("images")) {
    dir.create("images")
  }
  
  # Use R graphics to create an SVG file
  svg("images/vax_logo.svg", width = 6, height = 3)
  
  # Set up the plot
  par(mar = c(0,0,0,0))
  plot(0, 0, type = "n", xlim = c(0, 6), ylim = c(0, 3), axes = FALSE, xlab = "", ylab = "")
  
  # Draw five interlocking circles (Olympic rings style)
  symbols(c(1, 2.2, 3.4, 4.6, 5.8), c(1.5, 1.5, 1.5, 1.5, 1.5), 
          circles = rep(0.8, 5), inches = FALSE, add = TRUE, 
          fg = c("#3283C8", "#F1CF5B", "#53A842", "#E42526", "#28AAE1"), 
          bg = NA, lwd = 10)
  
  # Add text labels inside each circle
  text(c(1, 2.2, 3.4, 4.6, 5.8), c(1.5, 1.5, 1.5, 1.5, 1.5), 
       labels = c("V", "A", "X", "S", "G"), 
       col = "white", font = 2, cex = 1.2)
  
  dev.off()
}

# Try to create the logo (only if it doesn't exist)
if (!file.exists("images/vax_logo.svg")) {
  tryCatch({
    create_vax_logo()
  }, error = function(e) {
    warning("Could not create logo: ", e$message)
  })
}

# Read in the Wellcome data
full_df <- read_excel("~/Documents/STAT 541/wgm2018-dataset-crosstabs-all-countries.xlsx", sheet = 'Full dataset')
data_dict <- read_excel("~/Documents/STAT 541/wgm2018-dataset-crosstabs-all-countries.xlsx", sheet = 'Data dictionary')

# --- 1. Common Processing: Create Mappings ---

# Define a robust function to parse mappings from the data dictionary
parse_mapping <- function(mapping_raw) {
  # Check if mapping_raw is NA or empty
  if (is.na(mapping_raw) || nchar(trimws(mapping_raw)) == 0) {
    warning("Received empty or NA mapping string.")
    return(tibble(code = integer(), name = character()))
  }

  # Split by comma, potentially handling extra spaces
  mapping_list <- strsplit(mapping_raw, ",")[[1]]
  # Keep only elements containing '=', trim whitespace
  mapping_list <- trimws(mapping_list[grepl("=", mapping_list)])

  # Split each element by '='
  split_list <- strsplit(mapping_list, "=", fixed = TRUE)

  # Create a tibble, handling potential errors during conversion
  mapping_df <- tryCatch({
    tibble(
      code_raw = sapply(split_list, `[`, 1),
      name = sapply(split_list, `[`, 2)
    ) %>%
      # Attempt to convert code to integer, coercing errors to NA
      mutate(code = suppressWarnings(as.integer(trimws(code_raw))),
             name = trimws(name)) %>%
      # Filter out rows where code conversion failed or name is missing
      filter(!is.na(code), !is.na(name), name != "") %>%
      select(code, name)
  }, error = function(e) {
    warning("Error parsing mapping: ", e$message)
    return(tibble(code = integer(), name = character()))
  })

  return(mapping_df)
}

# Create Country Mapping Data Frame
country_mapping_raw <- data_dict %>%
  filter(`Variable Name` == "WP5") %>%
  pull(`Variable Type & Codes*`)
# Keep original mapping column names for joining ease, rename later if needed
country_mapping_df <- parse_mapping(country_mapping_raw) # Columns: code, name

# Create Region Mapping Data Frame
region_mapping_raw <- data_dict %>%
  filter(`Variable Name` == "Regions_Report") %>%
  pull(`Variable Type & Codes*`)
# Keep original mapping column names for joining ease
region_mapping_df <- parse_mapping(region_mapping_raw) # Columns: code, name


# --- 2. Analysis 1: Vaccine Safety Perception (Q25) ---

vaccine_safety_summary <- full_df %>%
  # Select necessary columns including the weight 'wgt'
  select(WP5, Regions_Report, wgt, Q25) %>%
  # Filter out missing or "Don't know" responses for Q25
  filter(!is.na(Q25) & Q25 != 99) %>%
  # Create binary indicator: 1 if they agree/strongly agree vaccines are safe
  mutate(vaccine_safe = ifelse(Q25 %in% c(1, 2), 1, 0)) %>%
  # Join with mapping tables *before* summarizing
  # Use suffixes to handle duplicated 'name' columns from mappings
  left_join(country_mapping_df, by = c("WP5" = "code"), suffix = c("", ".country")) %>%
  left_join(region_mapping_df, by = c("Regions_Report" = "code"), suffix = c(".country", ".region")) %>%
  # Filter out rows where mapping might have failed (optional but good practice)
  filter(!is.na(name.country), !is.na(name.region)) %>%
  # Group by the mapped names (which are now name.country and name.region)
  group_by(name.country, name.region) %>%
  # Calculate the weighted percentage, name it as in the original final df
  summarize(
    `% Believe Vaccines Safe` = sum(vaccine_safe * wgt, na.rm = TRUE) / sum(wgt, na.rm = TRUE) * 100,
    .groups = 'drop' # Drop grouping structure after summarizing
  ) %>%
  # Rename the grouped columns to match the original 'final_df' names
  rename(
      Country = name.country,
      Region = name.region
  ) %>%
  # Select only the final desired columns in the desired order
  select(Country, Region, `% Believe Vaccines Safe`)

# --- 3. Analysis 2: Child Vaccination Status (Q28) ---

vaccination_status_summary <- full_df %>%
  # Select necessary columns (Weight 'wgt' wasn't used in original code for this part)
  select(WP5, Q28) %>%
  # Keep only valid responses (1 = vaccinated, 2 = not vaccinated)
  filter(Q28 %in% c(1, 2)) %>%
  # Join with country mapping *before* summarizing
  left_join(country_mapping_df, by = c("WP5" = "code")) %>% # Adds 'name' column for country
  # Filter out rows where country mapping failed
  filter(!is.na(name)) %>%
  # Create binary indicator for vaccinated status
  mutate(Vaccinated = if_else(Q28 == 1, 1, 0)) %>%
  # Group by the mapped Country name (column 'name')
  group_by(name) %>%
  # Summarize counts and calculate proportion (unweighted as per original code)
  # Names generated here match the original 'country_summary'
  summarize(
    N_Respondents = n(),
    N_Vaccinated = sum(Vaccinated),
    Proportion_Vaccinated = N_Vaccinated / N_Respondents,
    .groups = "drop"
  ) %>%
  # Rename the 'name' column to 'Country' to match original variable name
  rename(Country = name) %>%
  # Apply specific country name adjustments as in the original code
  mutate(Country = case_when(
    Country == "Bosnia and Herzegovina" ~ "Bosnia and Herzegovina", # No change needed, but included for completeness
    Country == "Czech Republic" ~ "Czechia",
    Country == "Dominican Republic" ~ "Dominican Republic", # No change
    Country == "Eswatini" ~ "Eswatini", # No change
    Country == "Ivory Coast" ~ "Côte d'Ivoire",
    Country == "Macedonia" ~ "North Macedonia",
    Country == "Northern Cyprus" ~ "Cyprus", # Note: Original comment maintained
    Country == "Palestinian Territories" ~ "Palestine",
    Country == "Republic of Congo" ~ "Republic of the Congo",
    Country == "The Gambia" ~ "Gambia",
    Country == "United States" ~ "United States of America",
    TRUE ~ Country # Keep all other names as they are
  ))


# --- 4. Analysis 3: View of Science (ViewOfScience) ---

science_view_summary <- full_df %>%
  # Select necessary columns (Weight 'wgt' wasn't used in original code)
  select(WP5, ViewOfScience) %>%
  # Keep only valid responses (assuming 1-4 are valid categories)
  filter(ViewOfScience %in% c(1, 2, 3, 4)) %>%
  # Join with country mapping *before* summarizing
  left_join(country_mapping_df, by = c("WP5" = "code")) %>% # Adds 'name' column for country
  # Filter out rows where country mapping failed
  filter(!is.na(name)) %>%
  # Create readable labels for the science view categories
  mutate(
    ViewOfScience_Label = case_when(
      ViewOfScience == 1 ~ "Enthusiast",
      ViewOfScience == 2 ~ "Included",
      ViewOfScience == 3 ~ "Excluded",
      ViewOfScience == 4 ~ "Sceptic",
      TRUE ~ NA_character_ # Should not happen due to filter, but safe practice
    )
  ) %>%
  # Filter out any rows where label creation failed (optional)
  filter(!is.na(ViewOfScience_Label)) %>%
  # Group by Country (column 'name') and the created label
  group_by(name, ViewOfScience_Label) %>%
  # Count occurrences for each group (unweighted as per original code)
  # Name 'N' matches original
  summarize(N = n(), .groups = "drop") %>%
   # Rename the country name column *before* the next grouping
  rename(Country = name) %>%
  # Group again just by Country to calculate proportions within each country
  group_by(Country) %>%
  # Calculate the proportion of each view within the country
  # Name 'Proportion' matches original
  mutate(Proportion = N / sum(N)) %>%
  # Ungroup for a clean final tibble
  ungroup()

# Add broader region mapping for better visualization
region_mapping <- c(
  "Northern America" = "Americas",
  "South America" = "Americas",
  "Central America and Mexico" = "Americas",
  "Northern Europe" = "Europe",
  "Western Europe" = "Europe",
  "Southern Europe" = "Europe",
  "Eastern Europe" = "Europe",
  "North Africa" = "Middle East & North Africa",
  "Middle East" = "Middle East & North Africa",
  "South Asia" = "Asia",
  "Southeast Asia" = "Asia",
  "East Asia" = "Asia",
  "Western Africa" = "Sub-Saharan Africa",
  "Eastern Africa" = "Sub-Saharan Africa",
  "Southern Africa" = "Sub-Saharan Africa",
  "Central Africa" = "Sub-Saharan Africa",
  "Aus/NZ" = "Oceania",
  "Central Asia" = "Former Soviet Union",
  "Not assigned" = "Other"
)

# Add broader region to the dataset
vaccine_safety_summary <- vaccine_safety_summary %>%
  mutate(BroadRegion = region_mapping[Region])

# Calculate global statistics for the info cards
# 1. Function to create summary cards
create_summary_stats <- function() {
  # Calculate global statistics
  global_vax_belief <- mean(vaccine_safety_summary$`% Believe Vaccines Safe`, na.rm = TRUE)
  
  global_vax_coverage <- weighted.mean(
    vaccination_status_summary$Proportion_Vaccinated,
    vaccination_status_summary$N_Respondents,
    na.rm = TRUE
  ) * 100
  
  # Count countries with high/low vaccine confidence
  high_confidence_count <- sum(vaccine_safety_summary$`% Believe Vaccines Safe` > 90, na.rm = TRUE)
  low_confidence_count <- sum(vaccine_safety_summary$`% Believe Vaccines Safe` < 50, na.rm = TRUE)
  
  # Calculate global science perception stats
  science_global <- science_view_summary %>%
    group_by(ViewOfScience_Label) %>%
    summarize(
      total_n = sum(N),
      proportion = sum(N) / sum(science_view_summary$N),
      .groups = "drop"
    )
  
  # Return the values
  return(list(
    global_vax_belief = global_vax_belief,
    global_vax_coverage = global_vax_coverage,
    high_confidence_count = high_confidence_count,
    low_confidence_count = low_confidence_count,
    total_countries = nrow(vaccine_safety_summary),
    science_global = science_global
  ))
}

# Get stats for cards
card_stats <- create_summary_stats()

# 2. Function to create regional comparison plot for vaccine attitudes
create_regional_breakdown <- function() {
  # Group by region and calculate statistics
  region_breakdown <- vaccine_safety_summary %>%
    group_by(BroadRegion) %>%
    summarize(
      mean_belief = mean(`% Believe Vaccines Safe`, na.rm = TRUE),
      min_belief = min(`% Believe Vaccines Safe`, na.rm = TRUE),
      max_belief = max(`% Believe Vaccines Safe`, na.rm = TRUE),
      range_belief = max_belief - min_belief,
      count = n(),
      .groups = "drop"
    ) %>%
    arrange(desc(mean_belief))
  
  # Create the plot
  p <- ggplot(region_breakdown, aes(x = reorder(BroadRegion, mean_belief), y = mean_belief)) +
    geom_bar(stat = "identity", fill = "#3498DB", alpha = 0.7) +
    geom_errorbar(aes(ymin = min_belief, ymax = max_belief), width = 0.2, color = "#E74C3C") +
    geom_text(aes(label = sprintf("%.1f%%", mean_belief)), hjust = -0.1) +
    coord_flip() +
    theme_minimal() +
    labs(
      title = "Vaccine Safety Belief by Region",
      subtitle = "Average percentage with range (min-max)",
      x = NULL,
      y = "% Believe Vaccines are Safe"
    ) +
    theme(
      plot.title = element_text(face = "bold"),
      panel.grid.minor = element_blank()
    )
  
  return(p)
}

# 3. Function to create top/bottom countries visualization
create_topbottom_countries <- function() {
  # Get top and bottom 10 countries
  top10 <- vaccine_safety_summary %>%
    arrange(desc(`% Believe Vaccines Safe`)) %>%
    head(10)
  
  bottom10 <- vaccine_safety_summary %>%
    arrange(`% Believe Vaccines Safe`) %>%
    head(10)
  
  # Combine and add rank factor
  combined <- bind_rows(
    top10 %>% mutate(group = "Highest 10"),
    bottom10 %>% mutate(group = "Lowest 10")
  ) %>%
    group_by(group) %>%
    mutate(rank = row_number()) %>%
    ungroup()
  
  # Create the plot
  p <- ggplot(combined, aes(x = reorder(Country, `% Believe Vaccines Safe`), 
                          y = `% Believe Vaccines Safe`,
                          fill = group)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = sprintf("%.1f%%", `% Believe Vaccines Safe`)), 
            hjust = -0.1, size = 3) +
    facet_wrap(~group, scales = "free_y") +
    coord_flip() +
    scale_fill_manual(values = c("Highest 10" = "#27AE60", "Lowest 10" = "#E74C3C")) +
    theme_minimal() +
    theme(
      legend.position = "none",
      plot.title = element_text(face = "bold"),
      strip.text = element_text(face = "bold", size = 10),
      panel.grid.minor = element_blank()
    ) +
    labs(
      title = "Countries with Highest and Lowest Vaccine Confidence",
      x = NULL,
      y = "% Believe Vaccines are Safe"
    )
  
  return(p)
}

# 4. Function to create regional science perception comparison
create_region_science_comparison <- function() {
  # Join science view data with regions
  science_by_region <- science_view_summary %>%
    left_join(select(vaccine_safety_summary, Country, BroadRegion), by = "Country") %>%
    filter(!is.na(BroadRegion))
  
  # Group by region and ViewOfScience
  region_science <- science_by_region %>%
    group_by(BroadRegion, ViewOfScience_Label) %>%
    summarize(
      Total_N = sum(N),
      Weighted_Proportion = sum(N * Proportion) / sum(N),
      .groups = "drop"
    ) %>%
    # Add category colors
    mutate(
      color = case_when(
        ViewOfScience_Label == "Enthusiast" ~ "#FFD166",
        ViewOfScience_Label == "Included" ~ "#06D6A0", 
        ViewOfScience_Label == "Excluded" ~ "#73D2DE",
        ViewOfScience_Label == "Sceptic" ~ "#FFC5D3",
        TRUE ~ "#CCCCCC"
      )
    )
  
  # Create plot
  p <- ggplot(region_science, 
            aes(x = BroadRegion, y = Weighted_Proportion, 
                fill = ViewOfScience_Label)) +
    geom_bar(stat = "identity", position = "stack") +
    coord_flip() +
    scale_fill_manual(values = c(
      "Enthusiast" = "#FFD166", 
      "Included" = "#06D6A0", 
      "Excluded" = "#73D2DE", 
      "Sceptic" = "#FFC5D3"
    )) +
    scale_y_continuous(labels = scales::percent) +
    theme_minimal() +
    theme(
      legend.title = element_blank(),
      legend.position = "bottom",
      plot.title = element_text(face = "bold"),
      panel.grid.minor = element_blank()
    ) +
    labs(
      title = "Science Perception Categories by Region",
      x = NULL,
      y = "Proportion"
    )
  
  return(p)
}

# 5. Function to create correlation analysis between vaccine safety and science views
create_correlation_plot <- function() {
  # Prepare data
  # Calculate proportion of enthusiasts by country
  enthusiast_by_country <- science_view_summary %>%
    filter(ViewOfScience_Label == "Enthusiast") %>%
    select(Country, Proportion) %>%
    rename(Enthusiast_Proportion = Proportion)
  
  # Join with vaccine safety data
  correlation_data <- vaccine_safety_summary %>%
    left_join(enthusiast_by_country, by = "Country") %>%
    filter(!is.na(Enthusiast_Proportion))
  
  # Calculate correlation
  corr_value <- cor(correlation_data$`% Believe Vaccines Safe`, 
                  correlation_data$Enthusiast_Proportion,
                  use = "complete.obs")
  
  # Create scatter plot
  p <- ggplot(correlation_data, 
            aes(x = `% Believe Vaccines Safe`, 
                y = Enthusiast_Proportion * 100,
                color = BroadRegion)) +
    geom_point(size = 3, alpha = 0.7) +
    geom_smooth(method = "lm", color = "black", se = TRUE, alpha = 0.2) +
    scale_color_brewer(palette = "Set1") +
    theme_minimal() +
    theme(
      legend.title = element_blank(),
      plot.title = element_text(face = "bold"),
      panel.grid.minor = element_blank()
    ) +
    labs(
      title = "Correlation between Vaccine Confidence and Science Enthusiasm",
      subtitle = paste("Correlation coefficient:", round(corr_value, 2)),
      x = "% Believe Vaccines are Safe",
      y = "% Science Enthusiasts"
    )
  
  return(p)
}
```

# Vaccination Insights {orientation="columns" nav-tab="active"}

## Column {width=25%}

### Global Vaccine Confidence{height=50%}

::: {.valuebox icon="award" color="#3498db"}
Global Average Belief in Vaccine Safety:

`{r} paste0(round(card_stats$global_vax_belief, 1), "%")`
:::

### Vaccination Coverage{height=50%}

::: {.valuebox icon="clipboard-check" color="#2ecc71"}
Global Average Vaccination Rate:

`{r} paste0(round(card_stats$global_vax_coverage, 1), "%")`
:::

## Column {.tabset width=75%}

### Vaccine Safety Perception 

```{r}
#| label: new-and-improved-visualization
#| fig-width: 10
#| fig-height: 7
#| fig-align: center

# Calculate region medians
region_stats <- vaccine_safety_summary %>%
  group_by(BroadRegion) %>%
  summarise(
    median_value = median(`% Believe Vaccines Safe`),
    count = n()
  ) %>%
  arrange(desc(median_value))

# Get the order of regions by median value (descending)
region_order <- region_stats$BroadRegion

# Identify global extremes (top 5 and bottom 5)
top5 <- vaccine_safety_summary %>%
  arrange(desc(`% Believe Vaccines Safe`)) %>%
  head(5)

bottom5 <- vaccine_safety_summary %>%
  arrange(`% Believe Vaccines Safe`) %>%
  head(5)

# Combine top and bottom into extremes
extremes <- bind_rows(top5, bottom5)

# Create a new dataset with highlight information
vaccine_safety_summary <- vaccine_safety_summary %>%
  mutate(
    highlighted = Country %in% c(top5$Country, bottom5$Country),
    point_size = ifelse(highlighted, 3, 1.5),
    point_alpha = ifelse(highlighted, 1, 0.7)
  )

# Define region colors
region_colors <- c(
  "Asia" = "#D35400",
  "Middle East & North Africa" = "#3498DB",
  "Sub-Saharan Africa" = "#27AE60",
  "Americas" = "#E67E22",
  "Europe" = "#7F8C8D",
  "Oceania" = "#16A085",
  "Former Soviet Union" = "#8E44AD",
  "Other" = "#95A5A6"
)

# Create the plot
p <- ggplot() +
  # Add horizontal lines for each region
  geom_segment(
    data = region_stats,
    aes(
      x = 30, 
      xend = 100,
      y = factor(BroadRegion, levels = region_order),
      yend = factor(BroadRegion, levels = region_order),
      color = BroadRegion
    ),
    alpha = 0.3,
    size = 2
  ) +
  # Add points for each country
  geom_point(
    data = vaccine_safety_summary,
    aes(
      x = `% Believe Vaccines Safe`,
      y = factor(BroadRegion, levels = region_order),
      color = BroadRegion,
      size = point_size,
      alpha = point_alpha
    )
  ) +
  # Add vertical lines for region medians
  geom_segment(
    data = region_stats,
    aes(
      x = median_value,
      xend = median_value,
      y = as.numeric(factor(BroadRegion, levels = region_order)) - 0.2,
      yend = as.numeric(factor(BroadRegion, levels = region_order)) + 0.2,
      color = BroadRegion
    ),
    size = 1.5
  ) +
  # Add median value labels
  geom_text(
    data = region_stats,
    aes(
      x = median_value,
      y = factor(BroadRegion, levels = region_order),
      color = BroadRegion,
      label = paste0("Median: ", round(median_value, 1))
    ),
    vjust = -1.5,
    size = 3.5
  ) +
  # Add country labels for extremes
  geom_text(
    data = extremes,
    aes(
      x = `% Believe Vaccines Safe`,
      y = factor(BroadRegion, levels = region_order),
      label = Country
    ),
    hjust = -0.2,
    vjust = 1.2,
    size = 3,
    check_overlap = TRUE
  ) +
  # Set scale and appearance
  scale_x_continuous(
    limits = c(30, 100),
    breaks = seq(30, 100, by = 10),
    labels = function(x) paste0(x, "%"),
    expand = expansion(mult = c(0.01, 0.1))
  ) +
  scale_color_manual(values = region_colors) +
  scale_size_identity() +
  scale_alpha_identity() +
  # Customize theme
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    plot.caption = element_text(hjust = 0.5, size = 8),
    axis.title.y = element_blank()
  ) +
  # Add labels
  labs(
    title = "Percentage of people who believe vaccines are safe",
    subtitle = "By global region (2018)",
    x = NULL,
    caption = "Source: Wellcome Global Monitor, part of the Gallup World Poll 2018"
  )

# Add a text annotation for global extremes
p <- p + 
  annotate(
    "text", 
    x = 30, 
    y = 0.55, 
    label = "Global extremes are highlighted with larger points", 
    hjust = 0, 
    size = 3.2,
    fontface = "italic"
  )

p
```


### Vaccination Status 


```{r}
#| label: second-improved-visualization
#| fig-width: 10
#| fig-height: 7

# 1. Get world map shapefile
world <- ne_countries(scale = "medium", returnclass = "sf")

# 2. Join the vaccination data to the world map
world_vax <- world %>%
  left_join(vaccination_status_summary, by = c("name" = "Country"))

# 3. Create popup text
world_vax <- world_vax %>%
  mutate(
    popup_text = paste0(
      "<strong>Country: </strong>", name, "<br>",
      "<strong>Percent Vaccinated: </strong>", 
      ifelse(!is.na(Proportion_Vaccinated), scales::percent(Proportion_Vaccinated, accuracy = 0.1), "No data")
    )
  )

# Create the color palette function explicitly
# We define the domain so the palette function knows the range of values
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = world_vax$Proportion_Vaccinated,
  na.color = "#CCCCCC"
)

# 4. Create Leaflet map - Single World, Zoomed In Slightly
leaflet(data = world_vax) %>%
  addProviderTiles(
    providers$CartoDB.Positron,
    options = providerTileOptions(noWrap = TRUE) # Add this option to prevent repeating world
   ) %>%
  setView(lng = 0, lat = 30, zoom = 2) %>% # Set initial center and zoom level
  addPolygons(
    fillColor = ~pal(Proportion_Vaccinated), # Use the pre-defined palette function
    weight = 1,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 3,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE
    ),
    label = lapply(world_vax$popup_text, htmltools::HTML),
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto"
    )
  ) %>%
  # Add the legend to the map
  addLegend(
    pal = pal,                       # Use the same palette function
    values = ~Proportion_Vaccinated, # Use the data values the colors are based on
    opacity = 0.7,                   # Match the fill opacity used in addPolygons
    title = "Percent Vaccinated", # Set a title for the legend
    position = "bottomright"         # Specify legend position
  )
```

# Science Attitudes {orientation="columns"}

## Column {width=25%}

### Science Enthusiasts{height=33%}

```{r}
#| label: science-enthusiast-card

# Calculate global percentage of enthusiasts
enthusiast_pct <- card_stats$science_global %>%
  filter(ViewOfScience_Label == "Enthusiast") %>%
  pull(proportion) * 100
```

::: {.valuebox icon="book-fill" color="#f39c12"}
Global Science Enthusiasts:

`{r} paste0(round(enthusiast_pct, 1), "%")`
:::

### Science Skeptics{height=33%}

```{r}
#| label: science-skeptic-card

# Calculate global percentage of skeptics
skeptic_pct <- card_stats$science_global %>%
  filter(ViewOfScience_Label == "Sceptic") %>%
  pull(proportion) * 100
```

::: {.valuebox icon="question-circle" color="#e74c3c"}
Global Science Skeptics:

`{r} paste0(round(skeptic_pct, 1), "%")`
:::

### Correlation{height=33%}

```{r}
#| label: correlation-card

# Calculate correlation between vaccine confidence and science enthusiasm
enthusiast_by_country <- science_view_summary %>%
  filter(ViewOfScience_Label == "Enthusiast") %>%
  select(Country, Proportion) %>%
  rename(Enthusiast_Proportion = Proportion)

# Join with vaccine safety data
correlation_data <- vaccine_safety_summary %>%
  left_join(enthusiast_by_country, by = "Country") %>%
  filter(!is.na(Enthusiast_Proportion))

# Calculate correlation
corr_value <- cor(correlation_data$`% Believe Vaccines Safe`, 
                correlation_data$Enthusiast_Proportion,
                use = "complete.obs")
```

::: {.valuebox icon="graph-up" color="#3498db"}
Correlation: Vaccine Confidence & Science Enthusiasm

`{r} round(corr_value, 2)`
:::

## Column {.tabset width=75%}

### Science Perception Categories {.tabset}

#### Global Overview

```{r}
#| label: third-improved-visualization
#| fig-width: 10
#| fig-height: 12
#| fig-align: center

# Load necessary libraries (already loaded in setup)
# Calculate global averages for each category
global_stats <- science_view_summary %>%
  group_by(ViewOfScience_Label) %>%
  summarize(
    total_n = sum(N),
    weighted_prop = sum(N * Proportion) / sum(N)
  )

# Define quadrant positions and characteristics for each category
categories <- tibble(
  ViewOfScience_Label = c("Enthusiast", "Included", "Excluded", "Sceptic"),
  x_pos = c(0.75, 0.75, 0.25, 0.25), # x position in quadrant (right/left)
  y_pos = c(0.25, 0.75, 0.25, 0.75), # y position in quadrant (top/bottom)
  description = c(
    "Believe science benefits both society and individuals",
    "Believe science benefits individuals but not necessarily society",
    "Believe science benefits society but not individuals in their lives",
    "Don't see significant benefits from science at either level"
  ),
  color = c("#FFD166", "#06D6A0", "#73D2DE", "#FFC5D3"), # Gold, Green, Turquoise, Pink
  label_text = c("Enthusiasts", "The Included", "The Excluded", "Sceptics")
)

# Join the category info with the global stats
plot_data <- global_stats %>%
  left_join(categories, by = "ViewOfScience_Label") %>%
  mutate(percent_label = paste0(round(weighted_prop * 100, 1), "%"))

# Total sample size
total_sample <- sum(plot_data$total_n)

# PLOT 1: Main Quadrant Visualization
p1 <- ggplot() +
  # Add quadrant dividing lines
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "gray50", size = 1) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray50", size = 1) +
  
  # Add bubbles for each category
  geom_point(data = plot_data, 
             aes(x = x_pos, y = y_pos, size = weighted_prop, fill = ViewOfScience_Label),
             shape = 21, color = "white", alpha = 0.9) +
  
  # Add percentage labels inside bubbles
  geom_text(data = plot_data,
            aes(x = x_pos, y = y_pos, label = percent_label),
            color = "white", fontface = "bold", size = 4) +
  
  # Add category names above/below bubbles
  geom_text(data = plot_data,
            aes(x = x_pos, y = y_pos, label = label_text),
            vjust = -2.5, fontface = "bold", size = 3.5) +
  
  # Set scales
  scale_size_continuous(range = c(20, 40)) +
  scale_fill_manual(values = setNames(plot_data$color, plot_data$ViewOfScience_Label)) +
  
  # Set axis labels
  labs(
    x = "Perceived Benefits to Individuals",
    y = "Perceived Benefits to Society"
  ) +
  
  # Add axis direction indicators
  annotate("text", x = c(0.1, 0.9), y = c(0.05, 0.05), 
           label = c("Lower", "Higher"), size = 3.5) +
  annotate("text", x = c(0.05, 0.05), y = c(0.1, 0.9), 
           label = c("Lower", "Higher"), size = 3.5, angle = 90) +
  
  # Clean up the theme
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(size = 12, face = "bold"),
    plot.margin = margin(20, 20, 20, 20)
  ) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))

# PLOT 2: Category Explanations
explanation_df <- tibble(
  category = plot_data$label_text,
  description = plot_data$description,
  percentage = plot_data$percent_label,
  color = plot_data$color
)

p2 <- ggplot(explanation_df, aes(y = 4:1)) +
  geom_segment(aes(x = 0, xend = 0.9, yend = 4:1), color = "gray80") +
  geom_point(aes(x = 0, fill = category), size = 5, shape = 21, color = "white") +
  geom_text(aes(x = 0.1, label = paste0(category, " (", percentage, ")")), 
            hjust = 0, fontface = "bold", size = 3.5) +
  geom_text(aes(x = 0.1, label = description), 
            hjust = 0, vjust = 1.8, size = 3) +
  scale_fill_manual(values = setNames(explanation_df$color, explanation_df$category)) +
  labs(title = "Understanding the Categories") +
  theme_void() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 14, face = "bold", hjust = 0),
    plot.margin = margin(10, 10, 10, 10)
  ) +
  coord_cartesian(xlim = c(-0.1, 4))

# PLOT 3: Selected Countries Comparison
# Select diverse countries to compare
selected_countries <- c("United States", "China", "Germany", "India", "Brazil", "Nigeria")

# Filter for selected countries
country_data <- science_view_summary %>%
  filter(Country %in% selected_countries) %>%
  left_join(categories %>% select(ViewOfScience_Label, color, label_text), by = "ViewOfScience_Label") %>%
  mutate(label_short = gsub("s$", "", label_text))  # Shorter labels for the chart

p3 <- ggplot(country_data, aes(x = label_short, y = Proportion, fill = ViewOfScience_Label)) +
  geom_bar(stat = "identity", width = 0.7) +
  geom_text(aes(label = paste0(round(Proportion * 100, 0), "%")), 
            position = position_stack(vjust = 0.5), 
            color = "white", size = 2.8, fontface = "bold") +
  facet_wrap(~Country, nrow = 2) +
  scale_fill_manual(values = setNames(categories$color, categories$ViewOfScience_Label)) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  labs(
    title = "Country Comparisons",
    x = NULL,
    y = "Percentage"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    axis.text.y = element_text(size = 8),
    strip.text = element_text(face = "bold", size = 10),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5)
  )

# Combine plots using patchwork
combined_plot <- (p1) / (p2 + p3) +
  plot_layout(heights = c(1.2, 1)) +
  plot_annotation(
    title = "How People Feel About the Benefits of Science",
    subtitle = "Combined view of perceived benefits at personal and societal levels",
    caption = paste0("Global survey of 144 countries (n=", format(total_sample, big.mark=","), ")\n",
                    "Note: ~15% of people did not answer one or both questions and were not included in the analysis."),
    theme = theme(
      plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      plot.caption = element_text(size = 10, hjust = 1)
    )
  )

combined_plot
```

#### Regional Comparison

```{r}
#| label: region-science-comparison
#| fig-width: 10
#| fig-height: 7

# Create regional science comparison
create_region_science_comparison()
```

### Science & Vaccines {.tabset}

#### Correlation Analysis

```{r}
#| label: science-vaccine-correlation
#| fig-width: 10
#| fig-height: 7

# Create correlation plot
create_correlation_plot()
```

#### Country Explorer {.no-title}

```{r}
#| label: country-explorer
#| fig-width: 10
#| fig-height: 7

# Create an interactive scatterplot to explore countries
# Join data
explorer_data <- vaccine_safety_summary %>%
  left_join(
    science_view_summary %>% 
      filter(ViewOfScience_Label == "Enthusiast") %>%
      select(Country, Proportion) %>%
      rename(Enthusiast_Proportion = Proportion),
    by = "Country"
  ) %>%
  left_join(
    science_view_summary %>% 
      filter(ViewOfScience_Label == "Sceptic") %>%
      select(Country, Proportion) %>%
      rename(Skeptic_Proportion = Proportion),
    by = "Country"
  ) %>%
  filter(!is.na(Enthusiast_Proportion), !is.na(Skeptic_Proportion)) %>%
  mutate(
    hover_text = paste0(
      "Country: ", Country, "<br>",
      "Region: ", BroadRegion, "<br>",
      "Vaccine Safety Belief: ", round(`% Believe Vaccines Safe`, 1), "%<br>",
      "Science Enthusiasts: ", round(Enthusiast_Proportion * 100, 1), "%<br>",
      "Science Skeptics: ", round(Skeptic_Proportion * 100, 1), "%"
    )
  )

# Create interactive plot with plotly
plot_ly(
  data = explorer_data,
  x = ~`% Believe Vaccines Safe`,
  y = ~(Enthusiast_Proportion * 100),
  size = ~(Skeptic_Proportion * 100),
  sizes = c(5, 30),
  color = ~BroadRegion,
  colors = "Set1",
  text = ~hover_text,
  hoverinfo = "text",
  type = "scatter",
  mode = "markers"
) %>%
  layout(
    title = "Interactive Country Explorer: Vaccine Confidence vs. Science Attitudes",
    xaxis = list(title = "% Believe Vaccines are Safe"),
    yaxis = list(title = "% Science Enthusiasts"),
    hovermode = "closest",
    showlegend = FALSE
  )
```